\documentclass[11pt]{amsart}

\usepackage{macros}

\linespread{1.25}

\usepackage[final]{pdfpages}

\setcounter{tocdepth}{2}

\title{Lecture 4: dg Modules for dg Lie algebras}

\def\dgVect{{\sf Vect}^{\sf dg}}
\def\dgLie{{\sf Lie}^{\sf dg}}
\def\dgAss{{\sf Ass}^{\sf dg}}
\def\dgCAlg{{\sf CAlg}^{\sf dg}}
\def\dgLienil{{\sf Lie}^{\sf dg, nil}}
\def\Moduli{{\sf Moduli}}
\def\dgMod{{\sf Mod}^{\sf dg}}

\begin{document}
\maketitle

The enveloping algebra of a Lie algebra recognizes quasi-isomorphisms of dg Lie algebras in the following way.

\begin{lem}
A map of dg Lie algebras $f : \fg \to \fh$ is a quasi-isomorphism if and only if the induced map
\[
U(f) : U(\fg) \to U(\fh)
\]
is a quasi-isomorphism of associative dg algebras. 
\end{lem}
\begin{proof}
We consider the spectral sequences induced by the natural filtrations on the enveloping algebras of $\fg$ and $\fh$. 
The map $U(f)$ clearly preserves the natural filtration on the enveloping algebra and so it induces a map of spectral sequences. 
In particular, it induces a map at the $E_1$ page, which is simply the associated graded algebras
\[
{\rm Gr} \; U(f) : \Sym^*(\fg) \to \Sym^*(\fh) .
\] 
The map of spectral sequence converges to the map $U(f) : U(\fg) \to U(\fh)$. 

Now, it is an easy exercise to show that $\fg \to \fh$ is a quasi-isomorphism if and only if it induces a quasi-isomorphism of free commutative dg algebras $\Sym^*(\fg) \to \Sym^*(\fh)$. 
Thus, in one direction, if $\fg \to \fh$ is a quasi-isomorphism, we obtain a quasi-isomorphism at the $E_1$-page, and the result follows. 

In the other direction, if $U(\fg) \to U(\fh)$ is a quasi-isomorphism, then taking associated gradeds we obtain a quasi-isomorphism $\Sym^*(\fg) \to \Sym^*(\fh)$ and hence $\fg \to \fh$ is a quasi-isomorphism by the above remark. 
\end{proof}

\section{Modules and (co)homology}

\subsection{(dg) Modules}

Recall that to every dg vector space $V$ we can associated an associative dg algebra $\End_k(V)$ of endomorphisms. 

\begin{dfn}
Let $\fg$ be a dg Lie algebra.
A {\bf dg $\fg$-module} is a dg vector space $M$ together with a map of dg Lie algebras
\[
\rho : \fg \to \End_k(M) .
\]
A map of dg modules is defined in the obvious way. 
Let $\dgMod_{\fg}$ denote the category of dg $\fg$-modules. 
\end{dfn}

\begin{rmk}
The notion of a (left) dg module for an associative dg algebra is defined in the natural way.
By the universal property of the enveloping algebra, we see that a dg module for $\fg$ is equivalent to a (left) dg module for the associative dg algebra $U(\fg)$. 
\end{rmk}

Unlike the category of dg Lie algebras, the category of dg $\fg$-modules is an abelian category in the obvious way. 

As a matter of notation, when $M, N \in \dgMod_\fg$, we will write 
\[
M \tensor_k N \in \dgVect_k
\]
to be the tensor product as underlying dg vector spaces. 
While it is true that this tensor product still has the structure of a dg Lie algebra, we will not use it in what immediately follows. \footnote{In fact, the category $\dgMod_\fg$ is symmetric monoidal.}

%Also, there is the notion of tensor product. 
%
%\begin{dfn}
%Suppose $M,N$ are two dg $\fg$-modules.
%Define the tensor product dg $\fg$-module $M \tensor N$ to be the tensor product of underlying dg vector spaces with $\fg$-module structure given by
%\[
%\rho_{M} \tensor 1 + 1 \tensor \rho_N : \fg \to \End_k(M \tensor_k N) = \End_k(M) \tensor_k \End_k(N) .
%\]
%\end{dfn}

%\begin{rmk}
%If we think of $M,N$ as dg $U\fg$-modules, then 
%\[
%M \tensor N = M \tensor_{U\fg} N
%\]
%as $U \fg$-modules. 
%\end{rmk}

%This tensor product endows the category $\dgMod_{\fg}$ with the structure of a {\em symmetric monoidal category}. 

\subsection{(Co)Homology}

In what follows we only recount facets of Lie algebra homology/cohomology that we will use later on. 
For a more complete survey of these ideas we recommend the textbook reference \cite{FF}.

As for ordinary modules, we have a pair of functors
\[
\begin{array}{cccl}
(-)^\fg : & \dgMod_\fg & \to & \dgVect_k  \\
& M & \mapsto & M^\fg = \{m \in M \; | \; x \cdot m = m \; , \; \forall x \in \fg\} \\
(-)_{\fg} : & \dgMod_{\fg} & \to & \dgVect_k \\
& M & \mapsto & M / \fg \cdot M 
\end{array}
\]
called the invariants/coinvariants respectively. 

\begin{rmk}
Note that 
\[
M^\fg = {\rm Hom}_{U \fg}(k, M)
\]
and 
\[
M_\fg = k \tensor_{U \fg} M .
\] 
Here, we regard $M$ as a left dg $U(\fg)$-module and $k$ as a trivial (left and right) dg $U(\fg)$-module. 
\end{rmk}

\begin{lem}
Consider the functor
\[
{\rm triv}_\fg : \dgVect_k \to \dgMod_\fg
\]
that sends a dg vector space to the trivial $\fg$-module.
The functor $M \mapsto M^\fg$ is left adjoint to ${\rm triv}_\fg$. 
The functor $M \mapsto M_\fg$ is right adjoint to ${\rm triv}_\fg$. 
\end{lem}

As a consequence, taking invariants/coinvariants is left/right exact respectively. 
This motivates the following definition. 

\begin{dfn}
Let $\fg$ be a dg Lie algebra.
The {\bf Lie algebra homology} of $\fg$ is the left derived functor of coinvariants
\[
\begin{array}{cccc}
H_*(\fg ; -) : & \dgMod_\fg & \to & \Vect_k \\ 
& M & \mapsto & \LL_*(-)_\fg (M) \\ .
\end{array}
\]
Similarly, the {\bf Lie algebra cohomology} of $\fg$ is the right derived functor of invariants
\[
\begin{array}{cccc}
H^*(\fg ; -) : & \dgMod_\fg & \to & \Vect_k \\ 
& M & \mapsto & \RR_*(-)^\fg (M) \\ .
\end{array}
\]
\end{dfn}

\begin{rmk}
Using the Tor and Ext notation, we can write
\[
H^{\rm Lie}_*(\fg ; M) = {\rm Tor}^{U \fg}_* (k, M)
\]
and 
\[
H_{\rm Lie}^*(\fg ; M) = {\rm Ext}^*_{U \fg}(k, M) .
\] 
\end{rmk}

To compute the Lie algebra homology, for instance, one uses the usual trick for derived functors. 
By first finding a projective resolution, tensoring, then computing the cohomology. 
We proceed by finding a projective resolution of $U\fg$

\subsubsection{}

First, we sketch the following general construction for dg Lie algebras.
Given a dg Lie algebra $\fg$ define its {\em cone} to be the dg Lie algebra ${\rm Cone}(\fg)$ to be
\[
{\rm Cone}(\fg)^n = \fg^{n+1} \oplus \fg^n
\]
with differential
\[
\d_n = \left(\begin{array}{cc} \d_{\fg,n+1} & 0 \\ {\rm id}_{\fg^{n+1}}  & \d_{\fg, n} \end{array}\right) : {\rm Cone}(\fg)^{n} = \fg^{n+1} \oplus \fg^{n} \to \fg^{n+2} \oplus \fg^{n+1} = {\rm Cone}(\fg)^{n+1}
\]
and bracket
\[
[(x,y), (x',y')] = \left([x,y]_{\fg}, [x,y'] + [y,x']\right) .
\]

\begin{lem}
There is a natural map of dg Lie algebras
\[
\fg \hookrightarrow {\rm Cone}(\fg) .
\]
Furthermore, ${\rm Cone}(\fg)$ is acyclic.
\end{lem} 
\begin{proof}
The construction of the cone above is a special case of a general construction. 
Given any map of dg vector spaces
\[
f : V \to W
\]
we can construct the {\em cone} of $f$, ${\rm Cone}(f)$, as follows. 
For a textbook reference on these sorts of things, see \cite{Weibel}. 
The complex ${\rm Cone}(\fg)$ is the special case where $V = W = \fg$ and $f = {\rm id}$. 

As an underlying vector space
\[
{\rm Cone}(f) = V[1] \oplus W
\]
(thus ${\rm Cone}(f)^n = V^{n+1} \oplus W^n)$. 
The differential is
\[
\d_{{\rm Cone}} =  \left(\begin{array}{cc} \d_{V[1]} & 0 \\ f[1]  & \d_W \end{array}\right) .
\]

Observe that the cone of the identity morphism is acyclic.
An explicit nullhomotopy is given by the identity map itself. 
\end{proof}

\begin{rmk}
The reason for the terminology comes from topology. 
If $f : X \to Y$ is a map of topological spaces then it induces a map of singular cochains $f^* : {\rm C}_{\rm sing}^*(Y) \to {\rm C}_{\rm sing}^*(Y)$. 
Then, in the construction above, ${\rm Cone}(f^*)$ is equal to singular cochains of the {\em mapping cone} of $f$.
From this intuition, it is easy to see why the cone of the identity is contractible / acyclic. 
\end{rmk}

By functoriality of the enveloping functor, the associative algebra $U({\rm Cone}(\fg))$ is acyclic and hence a resolution for the trivial $U \fg$-module.
Thus, we have
\[
H^{\rm Lie}_*(\fg, M) = H^*\left(U({\rm Cone}(\fg)) \tensor_{U \fg} M \right) .
\] 

\begin{dfn}
The Chevalley-Eilenberg cochain complex computing Lie algebra {\em homology} is the dg vector space
\[
\clieu_*(\fg ; M) := U({\rm Cone}(\fg)) \tensor_{U \fg} M .
\] 
Its cohomology is precisely the Lie algebra homology of $M$.
\end{dfn}

\begin{rmk}
Explicitly, as a graded vector space, the CE complex is of the form
\begin{align*}
\clieu_*(\fg ; M) & = U({\rm Cone}(\fg)) \tensor_{U \fg} M  \\ & = \left(\Sym(\fg[1]) \tensor_k U(\fg)\right) \tensor_{U \fg} M \\ &= \Sym(\fg[1]) \tensor_k M .
\end{align*}
Tracing through these isomorphisms, one can deduce that the differential is
\begin{align*}
\d_{CE} (x_1,\ldots, x_n) & = \sum_{i = 1}^n (\pm) x_1 \cdots x_{i-1} (\d x_i) x_{i+1} \cdots x_n \\ & + \sum_{i < j} (\pm) x_1\cdots \Hat{x_i} \cdots x_{j-1} [x_i,x_j] x_{j+1} \cdots x_n .
\end{align*}
Thus, we can write the differential as a sum of two terms $$\d_{CE} = \d_{\fg} + \d_{\rm Lie}$$ where $\d_{\fg}$ is the differential induced from that on the dg Lie algebra $\fg$ by the Leibniz rule (this preserves Sym-degree) and $\d_{\rm Lie}$ encodes the Lie bracket (this decreases Sym-degree by one). 
\end{rmk}

\subsubsection{}

There is a completely analogous construction for Lie algebra cohomology. 

\begin{dfn}
The Chevalley-Eilenberg cochain complex computing Lie algebra {\em cohomology} is the dg vector space
\[
\clie^*(\fg ; M) := U({\rm Cone}(\fg)) \tensor_{U \fg} M .
\] 
Its cohomology is precisely the Lie algebra homology of $M$
\[
H_{\rm Lie}^*(\fg ; M) = H^*\left(\Hom_{U \fg}(U({\rm Cone}(\fg)), M)\right) .
\]
\end{dfn}

\begin{rmk}
One can identify $\clie^*(\fg ; M)$ with a complex of the form
\[
\clie^*(\fg ; M) = \left(\Sym(\fg^\vee[-1]) \tensor_k M , \d^{CE}\right) .
\]
Note that, when $M = k$ there is an identification 
\[
\clie^*(\fg ; k) = \left(\clieu_*(\fg ; k)\right)^\vee = \Hom_k(\clieu_*(\fg ;k), k) .
\] 
\end{rmk}

\begin{warn}
We abusively refer to $\clieu_*(\fg ; -)$ and $\clie^*(\fg; -)$ as the Chevalley-Eilenberg (CE) chain and cochain complexes, respectively.
Note that both are really {\em cochain complexes} (equivalently dg vector spaces) since their differentials increase grading degree.
We keep this terminology since they compute Lie algebra homology and cohomology, respectively. 
\end{warn}

\subsubsection{}

The CE complexes for homology and cohomology are functorial in the module input.
They both determine functors
\[
\begin{array}{cccc}
\clieu_*(\fg ; -) : & \dgMod_\fg & \to & \dgVect_k\\
\clie^*(\fg ; -) : & \dgMod_{\fg} & \to & \dgVect_k .
\end{array}
\]

We are interested in a different type of functoriality in the case that $M = k$, the trivial module. 
In this case, we write $\clieu_*(\fg) = \clieu(\fg;k)$ and similarly for cohomology.
A silly statement is that this trivial module is {\em universal} in the sense that it is a module for all dg Lie algebras.
Thus, we can contemplate the functoriality of homology/cohomology in the Lie algebra factor. 

\begin{lem}
The CE complex for homology/cohomology determine functors
\[
\begin{array}{cccc}
\clieu_*(-) : & \dgLie_k & \to & \dgVect_k\\
\clie^*(-) : & \dgLie_k & \to & \left(\dgVect_k\right)^{op} .
\end{array}
\]
\end{lem}

\subsubsection{}

There is a filtration on CE chains/cochains similar to the one we discussed for the enveloping algebra.

For chains, the filtration is increasing 
\[
k = F^0 \clieu_*(\fg) \hookrightarrow F^1 \clieu_*(\fg) \hookrightarrow \cdots
\]
defined by
\[
F^{k} \clieu_*(\fg) = \left(\Sym^{\leq k} (\fg[1]) , \d_{CE}\right) 
\] 
with the obvious inclusions. 
Note that the successive quotients are
\[
F^k / F^{k-1} \cong \left(\Sym^k(\fg) , \d_{\fg}\right)
\]
where $\d_{\fg}$ denotes the differential on the symmetric algebra induced from that on the dg Lie algebra $\fg$ by the Leibniz rule. 
The filtration is complete in the sense that $\clieu_*(\fg) \cong {\rm colim}_{k} F^k$. 

The filtration on cochains is defined similarly. 

\begin{lem}
Let $\fg,\fh$ be dg Lie algebras.
If $f : \fg \to \fh$ is a quasi-isomorphism of dg Lie algebras, then the induced maps
\[
\begin{array}{cccc}
\clieu_*(f) : & \clieu_*(\fg) & \to & \clieu_*(\fh)\\
\clie^*(\fg) : & \clie^*(\fh) & \to & \clie^*(\fg) 
\end{array}
\]
are quasi-isomorphisms. 
\end{lem}
\begin{proof}
We prove the statement for chains.
The proof for cochains is completely analogous. 
Note that the induced map $\clieu_*(f)$ preserves Sym-degree, and in particular, preserves the filtration $\clieu_*(f) : F^\bullet \clieu_*(\fg) \to F^\bullet \clieu_*(\fh)$. 

We prove the claim by induction. 
When $n = 0$ it is obvious. 
In general, consider the map of short exact sequences induced from $F^{n-1} \to F^n \to F^n / F^{n-1}$:
\[
\begin{tikzcd}
F^{n-1} \clieu_* (\fg) \ar[r] \ar[d, "f"] & \clieu_*(\fg) \ar[r] \ar[d,"f"] & \Sym^n(\fg [1]) \ar[d, "f"] \\
F^{n-1} \clieu_* (\fh) \ar[r] & \clieu_*(\fh) \ar[r] & \Sym^n(\fh [1]) .
\end{tikzcd}
\]
By assumption, the left vertical arrow is a quasi-isomorphism.
Also, since $k$ is characteristic zero, the map $\tensor^n f : \tensor^n \fg \to \tensor^n \fh$ is a quasi-isomorphism since $f$ is. 
Thus, the middle arrow is a quasi-isomorphism.
\end{proof}

\bibliographystyle{alpha}
%\bibliographystyle{spmpsci}  
\bibliography{def}


\end{document}