\documentclass[11pt]{amsart}

\usepackage{macros}

\linespread{1.25}

\usepackage[final]{pdfpages}

\setcounter{tocdepth}{2}

\def\Fun{{\sf Fun}}
\def\Top{{\sf Top}}
\def\colim{{\rm colim}\;}
\def\Sing{{\rm Sing}}
\def\Cat{{\sf Cat}}


\title{Lecture 13: Transferring model structures}

\begin{document}
\maketitle

This week we have two goals: 
\begin{itemize}
\item[(1)] recall the standard model structure on the category of dg vector spaces, and transfer it to model category structures on dg Lie algebras and commutative dg algebras;
\item[(2)] formulate a variant of the nerve construction that produces an $\infty$-category from an ordinary model category. 
This is how we will define the underlying $\infty$-categories (Lie, commutative algebras) that make up our derived enhancement of deformation theory. 
\end{itemize}

\section{Transferring model structures}

\begin{thm}
\label{thm: trans}
Suppose
\[
\begin{tikzcd}
\sC\ar[r,bend left,"F",""{name=A, below}] & \sD\ar[l,bend left,"G",""{name=B,above}] \ar[from=A, to=B, symbol=\dashv]
\end{tikzcd}
\]
is an adjoint pair and that $\sC$ is a cofbrantly generated model category. 
Suppose further that
\begin{itemize}
\item[(1)] the functor $G$ commutes with sequential colimits;
\item[(2)] a map in $\sD$ that has the LLP with respect to $G^{-1}({\rm Fib}_{\sC}) \subset {\rm Mor}(\sD)$ is in $G^{-1}(\sW)$. 
\end{itemize}
Then, $\sD$ has the structure of a cofibrantly generated model category where 
\begin{itemize}
\item the sets
\[
\{F(i) \; | \; i \in {\rm Cof}_\sC\} \;\; , \;\; \{F(j) \; | \; j \in {\rm Cof}_{\sC} \cap \sW_{\sC} \}
\]
generated the cofibrations and acyclic cofibrations respectively;
\item $f \in {\rm Mor}(\sD)$ is a fibration / weak equivalence if and only if $G(f)$ is a fibration / weak equivalence. 
\end{itemize}
\end{thm}
\begin{proof}(Idea)
For a full proof see Chapter II of \cite{GJ}.
We won't repeat the proof, but let me hint at how each condition is used. 

The core of the proof boils down to verifying the factorization axiom. 
That is, we we need to show that every map $f : X \to Y$ in $\sD$ can be factored as
\[
X \xto{i} Z \xto{p} Y
\]
where $i$ is a cofibration and $p$ is an acyclic fibration.

<<<<<<< HEAD
\begin{proof} (Sketch)
The main thing to verify is that every morphism $f : X \to Y$ in $\sD$ factors as
\[
X \xto{i} Z \xto{p} Y
\]
where $j$ is a cofibration and $p$ is an acyclic fibration. 

To construct it, one defines an inductive sequence of cofibrations
\[
X = Z_0 \xto{i_0} Z_1 \xto{i_1} \cdots \to Z_n \xto{i_n} Z_{n+1} \to \cdots
\]
together with maps $p_n : Z_n \to Y$ satisfying $p_n = p_{n+1} \circ i_n$. 
Define 
\[
Z = \underset{n}{\colim} Z_n .
\]
By construction this fits into a sequence $X \xto{i} Z \xto{p} Y$. 
It is immediate that $i$ is a cofibration, one must show that $p$ is an acyclic fibration. 

For this, we need to see that given any diagram
\[
\begin{tikzcd}
A \ar[r] \ar[d,"j"] & G(Z) \ar[d, "G(p)"] \\
B \ar[r] & G(Y) 
\end{tikzcd}
\]
such that $j$ is a cofibration, a lift $B \to G(Z)$ exists. 
Since {\em $G$ commutes with sequential colimits} we see this becomes equivalent to constructing an inductive set of lifts for each diagram
\[
\begin{tikzcd}
A \ar[r] \ar[d,"j"] & G(Z_n) \ar[d, "G(p_n)"] \\
B \ar[r] & G(Y) .
\end{tikzcd}
\]
Now, we can use the adjunction to turn this into a lifting problem
\[
\begin{tikzcd}
F(A) \ar[r] \ar[d,"F(j)"] & Z_n \ar[d, "p_n"] \\
F(B) \ar[r] & Y 
\end{tikzcd}
\]
which exists by definition. 
\end{proof}


We can restate condition (2) as follows. 
First, call $f \in {\rm Mor}(\sD)$ a fibration / weak equivalence if and only if $G(f)$ is a fibration / weak equivalence in $\sC$. 
Then, (2) requires that if a map in $\sD$ has the LLP with respect to all fibrations in $\sD$ then its a weak equivalence. 

This condition is quite hard to verify in practice, but thankfully there is an easier condition one can verify to ensure this holds. 

\begin{prop}\label{prop: easy}
Let $F,G$ be adjoint functors as above where $\sC$ is {\em any} model category. 
Call $f \in {\rm Mor}(\sD)$ a fibration / weak equivalence if and only if $G(f)$ is a fibration / weak equivalence in $\sC$. 
If
\begin{itemize}
\item[(a)] every object in $\sD$ is fibrant;
\item[(b)] every object in $\sD$ has a path object;
\end{itemize}
then a map that has the LLP with respect to every fibration is a weak equivalence.
\end{prop}

\begin{rmk}
The path object is the dual notion to a cylinder object.
Recall, if $X$ is an object of a model category, a path object is an object $PX$ together with maps $s : X \to PX$, $(p_0,p_1) : PX \to X$ such that $s$ is a weak equivalence, $p_0 \times p_1$ is a fibration, and the diagram
\[
\begin{tikzcd}
X \ar[r,"s"', "\simeq"]  \ar[dr, "1 \times 1"'] & P X \ar[d, "p_0 \times p_1"] \\
& X \times X .
\end{tikzcd}
\]
commutes.
\end{rmk}

\section{Transferring from $\dgVect$}

First of all, there is a natural model structure on the category of dg vector spaces given by the following:
\begin{itemize}
\item a map $f : V \to W$ in $\dgVect_k$ is a fibration if an only if $f : V^n \to W^n$ is surjective for each $n$;
\item a map $f : V \to W$ in $\dgVect_k$ is a cofibration if and only if $f : V^n \to W^n$ is injective for each $n$;
\item a map $f : V \to W$ in $\dgVect_k$ is a weak equivalence if and only if $f$ is a quasi-isomorphism. 
\end{itemize}

\begin{ex}
Show that this is a model structure on $\dgVect_k$.
What are the fibrant / cofibrant objects?
\end{ex}

\begin{thm}[\cite{Hov}]
This model structure is a cofibrantly generated model structure on $\dgVect_k$. 
\end{thm}

\begin{rmk}
Suppose that $W$ is an ordinary vector space.
Define the dg vector space $D^n(W)$ by 
\[
\fc^n (W) = W[-n] \xto{\id_W} W[-n-1] .
\]
We see that $\fc^n(W)$ is acyclic and that there is a natural inclusion of dg vector spaces $W[-n] \to \fc^n(W)$.
Note that $\fc^n(W)$ is precisely the cone of the map of dg vector spaces $\id : W[-n] \to W[-n]$. 

The set of all inclusions
\[
\{k[-n] \to \fc^n(k)\}_{n \in \ZZ}
\]
is a generating set for the cofibrations above. 
The set of maps
\[
\{0 \to \fc^n(k)\}_{n \in \ZZ}
\]
is a generating set for the acyclic cofibrations. 
\end{rmk}

\subsection{The model structure on $\dgCAlg$}
\def\oblv{{\rm oblv}}
\def\Tens{{\rm Tens}}

We use Theorem \ref{thm: trans} to induce a model structure on commutative dg algebras in the following way. 
Recall the adjunction
\[
\begin{tikzcd}
\dgVect_k \ar[r,bend left,"\Sym(-)",""{name=A, below}] & \dgCAlg_k \ar[l,bend left,"{\rm oblv}_C",""{name=B,above}] \ar[from=A, to=B, symbol=\dashv]
\end{tikzcd}
\]
Where $\Sym(-) : V \mapsto \Sym(V)$ is the symmetric algebra functor, and $\oblv_C$ is the forgetful functor. 

In order to apply Theorem \ref{thm: trans}, we need to verify that the following two conditions are satisfied:
\begin{itemize}
\item[(1)] The functor $\oblv_C$ preserves sequential colimits;
\item[(2)] Call a map $f : A \to B$ of commutative dg algebras is a fibration if $\oblv_C(f)$ is a fibration. 
Then, we need to show that if a map of commutative dg algebras has the LLP with respect to every fibration, then it is a quasi-isomorphism.
\end{itemize}

\subsubsection{}

Item (1) follows from the stronger result:

\begin{lem}
The forgetful functor 
\[
\oblv_C : \dgCAlg_k \to \dgVect_k
\]
creates filtered colimits.
\end{lem}

\begin{rmk}
A functor $G : \sC \to \sD$ {\em creates} colimits for a diagram $F : J \to \sC$ if whenever $\colim (G \circ F)$ exists, the colimit $\colim F$ exists and $G(\colim F) = \colim (G \circ F)$.
\end{rmk}

\begin{proof}
Let $F : J \to \dgCAlg_k$ be any filtered diagram and denote by $G$ the composition of functors
\[
J \xto{F} \dgCAlg_k \xto{\oblv_C} \dgVect_k .
\]
Let $A$ be the colimit of this functor (which exists as all colimits do in $\dgVect_k$), which we can express by
\[
A = \underset{j \in J}{\colim} G(j) .
\] 
As we show below, the key property of a filtered category is that the diagonal morphism $J \to J \times J$ is {\em final}. 

Using this, we have the following composition of dg vector spaces
\[
A \tensor_k A = \left(\underset{j \in J}{\colim} G(j)\right) \tensor_k \left(\underset{j' \in J}{\colim} G(j')\right) \cong \underset{(j,j') \in J\times J}{\colim} G(j) \tensor_k G(j') \xto{\cong} \underset{j \in J}{\colim} G(j) \tensor_k G(j) .
\]
We used finality of the diagonal functor $J \to J \times J$ in the last isomorphism. 

Now, for each $j$ we have the multiplication map $\mu_j : F(j) \tensor_k F(j) \to F(j)$.
Applying $G$ and taking the colimit we obtain a map
\[
\underset{j \in J}{\colim} G(j) \tensor_k G(j) \to \underset{j \in J}{\colim} G(j) = A .
\]
Thus, combining these two composition we obtain a map $\mu : A \tensor_k A \to A$, as desired. 
It is an exercise to show that this endows $(A, \mu)$ with the structure of a dg commutative algebra. 
\end{proof}

In the proof we have used a useful property of the diagonal functor $J \to J \times J$ of a filtered category $J$. 
Namely, that it is {\em final}. 

\begin{dfn}
A functor $F : J \to I$ is {\em final} if it satisfies the either of the following two equivalent properties:
\begin{itemize}
\item[(a)] 
For every object $i \in I$, the undercategory 
\[
F^{i /} =  
\]
is nonempty and connected. 
\item[(b)] for all categories $\sC$ and functors $G : I \to \sC$, the natural map
\[
\colim_{J} (G \circ F) \xto{\cong} \colim_{I} G
\]
is an isomorphism in $\sC$. 
\end{itemize}
\end{dfn}

\subsubsection{}

To show (2) we appeal to Proposition \ref{prop: easy}. 
The first condition in Proposition \ref{prop: easy} is immediate since every dg vector space is fibrant. 
So, all we need to show is the following. 

\begin{lem}
Every object in $\dgCAlg_k$ has a path object.
\end{lem}
\begin{proof}
Let $A$ be a commutative dg algebra. 
Define
\[
M = \bigoplus_{n \in \ZZ} \bigoplus_{(a_1,a_2) \in (A \times A)_n} \fc^n(k) .
\]
This looks complicated, but its just a huge acyclic object that we will construct the path space out of. 
For each $(a_1,a_2) \in (A \times A)_n$ we have a map of dg vector spaces
\[
p_{(a_1,a_2)} : \fc^n(k) \to A \times A
\]
that sends the generator in degree $n$, $1 \in \fc^n(k)^n = k$ to $(a_0,a_1)$, and the generator in degree $n+1$, $1 \in \fc^n(k)^{n+1}$ to $\d_{A \times A} (a_0, a_1)$. 
By the universal property of direct sums, the collection of all of these define a map
\[
p : M \to A \times A
\]
and by the universal property of the symmetric algebra we have a map of commutative algebras
\[
\Sym(p) : \Sym(M) \to A \times A .
\]
Note that $\Sym(p)$ is degreewise surjective by construction. 

In any symmetric algebra we have a unit $1 \in \Sym^0(M) = k \subset \Sym(M)$. 
Compiling these facts we obtain a commutative diagram
\[
\begin{tikzcd}
A \ar[r, "\id \tensor 1"] \ar[dr, "\id \times \id"'] & A \tensor_k \Sym(M) \ar[d, "(\id \times \id) \tensor \Sym(p)"] \\
& A \times A .
\end{tikzcd}
\]
By construction, $\Sym(M)$ is acyclic, so the top map is a quasi-isomorphism. 
Also, the vertical right map is a fibration since $\Sym(p)$ is. 
\end{proof}

To summarize, we have endowed $\dgCAlg_k$ with a cofibrantly generated model category structure with the properties:
\begin{itemize}
\item[(1)] A map $f : A \to B$ of commutative dg algebras is a weak equivalence if and only if $\oblv_C(f) : \oblv_C(A) \to \oblv_C(B)$ is a weak equivalence in $\dgVect_k$.
That is, a quasi-isomorphism. 
\item[(2)] A map $f : A \to B$ of commutative dg algebras is a fibration if and only if $\oblv_C(f) : \oblv_C(A) \to \oblv_C(B)$ is a fibration in $\dgVect_k$.
That is, a degreewise surjection. 
\item[(3)] The set of maps
\[
\{{\rm Tens}_L(k[n]) \to {\rm Tens}_L (\fc^n(k))\}_{n \in \ZZ}
\]
generate the cofibrations. 
\end{itemize}


\subsection{The model structure on $\dgLie$}

We can also utilize the transfer theorem to endow $\dgLie_k$ with a model structure by appealing to the adjunction
\[
\begin{tikzcd}
\dgVect_k \ar[r,bend left,"{\rm free}_L(-)",""{name=A, below}] & \dgLie_k \ar[l,bend left,"{\rm oblv}_L",""{name=B,above}] \ar[from=A, to=B, symbol=\dashv]
\end{tikzcd}
\]
between the free Lie algebra functor and the forgetful functor from Lie algebras to vector spaces. 

The proof is very similar, so we just read off the properties of this model structure on $\dgLie_k$. 

\begin{itemize}
\item[(1)] A map $f : \fg \to \fh$ of dg Lie algebras is a weak equivalence if and only if $\oblv_L(f) : \oblv_L(\fg) \to \oblv_L(\fh)$ is a weak equivalence in $\dgVect_k$.
That is, a quasi-isomorphism. 
\item[(2)] A map $f : \fg \to \fh$ of dg Lie algebras is a fibration if and only if $\oblv_L(f), \oblv_L(\fh)$ is a fibration in $\dgVect_k$.
That is, a degreewise surjection. 
\item[(3)] The set of maps
\[
\{{\rm free}_L(k[n]) \to {\rm free}_L (\fc^n(k))\}_{n \in \ZZ}
\]
generate the cofibrations. 
\end{itemize}





\end{document}
